Smart Spreading Factor Assignment for LoRaWANs

### 2019-01-01 paper.tex
cite

\section{Introduction}
\par In the last few years, number of Internet of Things applications increase exponentially. \cite{7721743} Recent developments on LPWAN technologies has great impact on IoT application number growth. LPWAN technologies offer solutions to some of the oldest wireless communication challenges. Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication methods (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time. [CITE] Cellular networks can provide long range and high data rate but they are complex and consume too much power. Besides, most of the IoT applications don't require high data rate. Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best. \cite{7815384} LPWAN technologies fill the technology gap between short range and cellular communication by providing low power and long range communication. LPWAN technologies basically sacrifice data rate to provide low power consumption.

\par There are several emerging LPWAN technologies. LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well known LPWAN technologies. LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost. [CITE] Both LoRa and Sigfox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate. LoRa has open standard MAC protocol called LoRaWAN. [CITE] LoRaWAN and Sigfox MAC protocols are based on pure ALOHA medium access. [CITE] LoRaWAN networks can be deployed as private network. However, Sigfox and NB-IoT supports only operator contracted deployments. [CITE] Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink. Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink. [CITE] However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload. [CITE] Sigfox maximum data rate is also very low 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively. [CITE]

\par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between receive sensitivity and transmission air time. [CITE] LoRa supports 6 different SF option. Simultaneous same SF transmissions are prone to collision however different SF transmissions in same channel are orthogonal to each other, thus SF assignment is crucial for overall network performance. [CITE] In this work, we introduce a custom made LoRa discrete event simulator to study LoRa SF assignment issue. We also proposed an SF assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using our simulator.

@@

\subsection{End node}
\par LoRaWAN end node (EN) is low power embedded device that communicates to gateways. LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C for providing LPWAN solution to different scenarios. Class A end nodes generates uplink transmission at any time and only receives a period of time after uplink transmission. Class B end nodes extend Class A by adding scheduled receive windows for downlink transmission. Receive window is synchronized using beacon packet transmitted by gateway. Class C end nodes extend Class A by keeping receive window open unless they are transmitting. This provides low latency donwlink communication but requires more power consumption. [CITE] In this paper, we focused on Class A end devices since Class A provides lowest power consumption.

@@

\section{LoRa}
\par LoRa is the name of the proprietary physical layer radio and chipset technology that provides wireless link solution for low power wide area networks. LoRa is designed and patented by Semtech Corporation. LoRa uses proprietary spread spectrum modulation technique that is derivative of Chirp Spread Spectrum (CSS). Chirp is a sinusoidal signal which its  frequency increases over time. The chirp frequency increases linearly and sweeps entire bandwidth. [CITE] We invite readers to refer to \cite{AN1200.22} for further details.

@@

\begin{table}
\centering
\caption{EU863-870 ISM Band LoRa Default Channels \cite{lorawan.regional.parameters}}
\label{table:max_tx_power}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Channel} & \textbf{Frequency (MHz)} & \textbf{Bandwidth (kHz)} & \textbf{Max EIRP (dBm)} \\ \hline
      1 &     868.1 &   125 &   14 \\ \hline
      2 &     868.3 &   125 &   14 \\ \hline
      3 &     868.5 &   125 &   14 \\

@@

\begin{table}
\centering
\caption{Gateway Sensitivity for different SFs \cite{SX1276}}
\label{table:gw_sf_sensitivity}
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
\multicolumn{2}{|c|}{\multirow{2}{*}{}} & \multicolumn{6}{c|}{\textbf{SF}} \\ \cline{3-8}
\multicolumn{2}{|c|}{}                  &    7 &    8 &    9 &   10 &   11 &   12 \\ \hline
\multirow{3}{*}{\textbf{BW (kHz)}}  & 125 & -123 & -126 & -129 & -132 & -133 & -136 \\ \cline{2-8}
                                    & 250 & -120 & -123 & -125 & -128 & -130 & -133 \\ \cline{2-8}
                                    & 500 & -116 & -119 & -122 & -125 & -128 & -130 \\ \hline
\end{tabular}
\end{table}

@@

%% \par \cite{7815384} \cite{7803607} \cite{7996384} \cite{8090518} \cite{s17061193} \cite{8267219} \cite{8430542} \cite{8319183} \cite{8480649} \cite{AN1200.22} \cite{Bor:2016:LLW:2988287.2989163} \cite{8406255} \cite{DBLP:journals/corr/abs-1802-10338} \cite{finnegan2018comparative}

>>>>

\section{Introduction}
\par In the last few years, number of Internet of Things applications increase exponentially \cite{7721743}. Recent developments on LPWAN technologies has great impact on IoT application number growth. LPWAN technologies offer solutions to some of the oldest wireless communication challenges. Traditional wireless communication methods such as cellular networks (e.g., 2G, 3G, LTE) and short-range communication methods (e.g., Bluetooth, WiFi, Zigbee) cannot provide low power and long range at the same time. Cellular networks can provide long range and high data rate but they are complex and consume too much power. Besides, most of the IoT applications don't require high data rate. Short-range communication methods can provide relatively low power consumption but their range is limited to a few hundred meters at best \cite{7815384}. LPWAN technologies fill the technology gap between short range and cellular communication by providing low power and long range communication. LPWAN technologies basically sacrifice data rate to provide low power consumption.

\par There are several emerging LPWAN technologies. LoRa, Sigfox, NB-IoT and LTE-M are commonly used, well known LPWAN technologies. LoRa and Sigfox use license free ISM frequency bands while NB-IoT and LTE-M use licensed frequency bands which brings extra cost \cite{7815384}. Both LoRa and Sigfox known for ultra low power consumption and resilient to interference while NB-IoT and LTE-M are promoted for higher data rate. LoRa has open standard MAC protocol called LoRaWAN. LoRaWAN and Sigfox MAC protocols are based on pure ALOHA medium access.  LoRaWAN networks can be deployed as private network. However, Sigfox and NB-IoT supports only operator contracted deployments \cite{7815384}. Number of messages that Sigfox end device can send in a day is limited to 140 packets for uplink and just 4 packets for downlink. Sigfox packet payload is limited to 12 bytes for uplink and 8 bytes for downlink. However LoRa supports up to 243 bytes payload and NB-IoT supports up to 1600 bytes payload. Sigfox maximum data rate is also very low 100 bps, on the other hand maximum data rate for LoRa and NB-IoT are 50 kbps and 200 kbps respectively \cite{7815384}.

\par LoRa can adjust data rate by spreading symbols within a fixed channel bandwidth. This enables tradeoff between receive sensitivity and transmission air time \cite{7803607}. LoRa supports 6 different SF option. Simultaneous same SF transmissions are prone to collision however different SF transmissions in same channel are orthogonal to each other, thus SF assignment is crucial for overall network performance. In this work, we introduce a custom made LoRa discrete event simulator to study LoRa SF assignment issue. We also proposed an SF assignment scheme to increase network performance with the help of machine learning methods and we verified our proposal using our simulator.

@@

\subsection{End node}
\par LoRaWAN end node (EN) is low power embedded device that communicates to gateways. LoRaWAN standard defines three classes for end devices which are Class A, Class B, and Class C for providing LPWAN solution to different scenarios. Class A end nodes generates uplink transmission at any time and only receives a period of time after uplink transmission. Class B end nodes extend Class A by adding scheduled receive windows for downlink transmission. Receive window is synchronized using beacon packet transmitted by gateway. Class C end nodes extend Class A by keeping receive window open unless they are transmitting. This provides low latency donwlink communication but requires more power consumption. In this paper, we focused on Class A end devices since Class A provides lowest power consumption.

@@

\section{LoRa}
\par LoRa is the name of the proprietary physical layer radio and chipset technology that provides wireless link solution for low power wide area networks. LoRa is designed and patented by Semtech Corporation. LoRa uses proprietary spread spectrum modulation technique that is derivative of Chirp Spread Spectrum (CSS). Chirp is a sinusoidal signal which its  frequency increases over time. The chirp frequency increases linearly and sweeps entire bandwidth. We invite readers to refer to \cite{AN1200.22} for further details.

@@

\begin{table}
\centering
\caption{EU863-870 ISM Band LoRa Default Channels \cite{lorawan.regional.parameters}.}
\label{table:max_tx_power}
\begin{tabular}{|c|c|c|c|}
\hline
\textbf{Channel} & \textbf{Frequency (MHz)} & \textbf{Bandwidth (kHz)} & \textbf{Max EIRP (dBm)} \\ \hline
      1 &     868.1 &   125 &   14 \\ \hline
      2 &     868.3 &   125 &   14 \\ \hline
      3 &     868.5 &   125 &   14 \\

@@

\begin{table}
\centering
\caption{Gateway Sensitivity for different SFs \cite{SX1276}.}
\label{table:gw_sf_sensitivity}
\begin{tabular}{|c|c|c|c|c|c|c|c|}
\hline
\multicolumn{2}{|c|}{\multirow{2}{*}{}} & \multicolumn{6}{c|}{\textbf{SF}} \\ \cline{3-8}
\multicolumn{2}{|c|}{}                  &    7 &    8 &    9 &   10 &   11 &   12 \\ \hline
\multirow{3}{*}{\textbf{BW (kHz)}}  & 125 & -123 & -126 & -129 & -132 & -133 & -136 \\ \cline{2-8}
                                    & 250 & -120 & -123 & -125 & -128 & -130 & -133 \\ \cline{2-8}
                                    & 500 & -116 & -119 & -122 & -125 & -128 & -130 \\ \hline
\end{tabular}
\end{table}

@@
